---
title: Clustering European Demography Data
output: html_document
---

### About

We use European demography data to test hierarchical clustering methods.

### Implementation

First, we load a special clustering library `cluster` although R already provides clustering in its base function set.

```{r message = FALSE, warning = FALSE}
library(cluster)
```

Furthermore, we load the sample data from our data repository online

```{r}
load(url("http://vault.data-r.com/data/demography.rda"))
```

```{r}
dim(demography)
names(demography)
demography$Country
head(demography)
```

Let's plot two parameters and select the fertility rate as well as the annual net migrationrate.

```{r echo = FALSE}
plot(demography$Fertilityrate, demography$Annualnetmigrationrate, 
     xlim=c(1,2.2), xlab="Fertility rate", ylab="Annual net migration rate")
textpos <- rep(4, 27)
textpos[8] <- textpos[10] <- textpos[20] <- textpos[21] <- 2
text(demography$Fertilityrate, demography$Annualnetmigrationrate, 
     demography$Country, pos=textpos, cex=0.7)
```

Now cluster the countries regarding these two attributes (fertility rate / annual net migrationrate) and plot the result.

```{r}
data <- na.omit(demography[, 1:4])
clustering <- agnes(data[, 3:4], stand = TRUE)
pltree(clustering, main='Hierarchical Clustering', sub="", xlab='EU Countries', labels=data$Code)
```

Transpose the data set and cluster: Demography-Variables

```{r}
data <- na.omit(demography)[, 3:19]
clustering <- agnes(t(data), stand = TRUE)
pltree(clustering, main="", sub="", xlab='Demography parameters')
```
