---
title: Association Analysis - Market Basket Analysis
output: html_document
---

### About

P. Giudici. Applied Data Mining. Wiley. 2003. Chapter 7:

- 75 days between January 2nd and April 21st, 2001.
- Southern Italy.
- 7301 loyalty cards.
- Average expenditure per transaction: EUR 28,27.
- Analysis limited to 20 categories (items), only food products.

### Implementation

First we load all required libraries. We need the library `arules` for mining association rules and `bit64` to work with long integer values properly.

```{r message = FALSE, warning = FALSE}
library(arules)
library(bit64)
```

Furthermore, we load the sample data from our data repository online

```{r}
load(url("http://vault.data-r.com/data/marketbasket.rda"))
```

```{r}
head(marketbasket)
```

First, create a special transaction set out of the the list of transactions and remove duplicates.

```{r}
write.csv2(marketbasket, file="transactions.txt", row.names = FALSE)
trans <- read.transactions("transactions.txt", format="single", sep=";", cols=c(1,2), rm.duplicates=TRUE)
```

Define minimum support and confidence:

- Support indicates how frequently an itemset appears in the data.
- Confidence indicates how often a rule was found to be true.

```{r}
support <- 0.2
confidence <- 0.8
```

Mine rules using the Apriori algorithm and check the top 30 rules sorted by Confidence.

```{r}
rules <- apriori(trans, parameter = list(supp = support, conf = confidence, target = "rules"))
inspect(head(rules, n = 30, by = "confidence"))
```

The lift of a rule is defined as the ratio of the observed support which is expected if X and Y were independent. The higher its value the more the two occurrences are dependent on one another, i.e. rules with a higher lift are potentially more useful for predictions.
