---
title: Movie Recommendation - Collaborative Filtering
output: html_document
---

### About

The dataset used in this case is the Movielens 100k set obtained from https://grouplens.org/datasets/movielens/100k/ which contains:

- 100.000 movie ratings
- 943 raters
- 1682 movies
- Ratings between 1 (worst) and 5 (best)

### Implementation

First we load all required libraries. We need the library `recommenderlab` for everything we need in regard to collaborative filtering and recommendation engines.

```{r message = FALSE, warning = FALSE}
library(recommenderlab)
```

Furthermore, we load movie data from our data repository online. See the Appendix at the end of this file to see how the original data was imported.

```{r}
load(url("http://vault.data-r.com/data/movierecommend.rda"))
```

and check the data

```{r}
dim(movie)
head(movie)
dim(data)
head(data)
```

First, build a (real-valued, 1-5) rating matrix and plot the ratings of the first 20 users on the first 20 movies

```{r}
ratings <- as(data[,1:3], "realRatingMatrix")
image(ratings[1:20,1:20])
```

Build a user-based collaborative filtering recommender with one line

```{r}
recommender <- Recommender(ratings, method="UBCF")
```

Try recommender, i.e. get a list of ids of recommended movies

```{r}
current.user <- 7

pre <- predict(recommender, current.user, data = ratings, n = 10)
as.numeric(unlist(as(pre, "list")))
```

and map the ids to move titles

```{r}
users.recommendations <- as.character(movie$movie.title[which(movie$movie.id %in% as.numeric(unlist(as(pre, "list"))))])
users.recommendations
```

Finally, we may take a closer list at all the movies the user rated, along with the ratings. Uncomment the line below to see the list

```{r}
users.ratings <- data.frame(movie$movie.title[data[which(data$user.id == current.user),2]], data[which(data$user.id == current.user),3])
names(users.ratings) <- c("Movie", "Rating")
# users.ratings
```

```{r include = FALSE}
### Appendix: Data Import

## The original data has been imported using the following code:
# 
# data <- read.delim("u.data", sep="\t", header=FALSE)
# names(data) <- c("user.id", "movie.id", "rating", "timestamp")
# 
# movie <- read.delim("u.item", sep="|", header=FALSE)
# names(movie) <- c("movie.id", "movie.title", "release.date", "video.release.date", "IMDb.URL", "unknown", "Action", "Adventure", "Animation", "Children", "Comedy", "Crime", "Documentary", "Drama", "Fantasy", "Film-Noir", "Horror", "Musical", "Mystery", "Romance", "Sci-Fi", "Thriller", "War", "Western")
# 
# save(data, movie, file="movierecommend.rda")
```
