---
title: Online Retailer Analysis
output: html_document
---

### About

Online Retail Data Set - a transnational data set which contains all the transactions occurring between 01/12/2010 and 09/12/2011 for a UK-based and registered non-store online retail. The company mainly sells unique all-occasion gifts. Many customers of the company are wholesalers.

https://archive.ics.uci.edu/ml/datasets/Online+Retail

- InvoiceNo: Invoice number. Nominal, a 6-digit integral number uniquely assigned to each transaction. If this code starts with letter 'c', it indicates a cancellation.
- StockCode: Product (item) code. Nominal, a 5-digit integral number uniquely assigned to each distinct product.
- Description: Product (item) name. Nominal.
- Quantity: The quantities of each product (item) per transaction. Numeric.
- InvoiceDate: Invice Date and time. Numeric, the day and time when each transaction was generated.
- UnitPrice: Unit price. Numeric, Product price per unit in sterling.
- CustomerID: Customer number. Nominal, a 5-digit integral number uniquely assigned to each customer.
- Country: Country name. Nominal, the name of the country where each customer resides.

#### Literature

- Daqing Chen, Sai Liang Sain, and Kun Guo, Data mining for the online retail industry: A case study of RFM model-based customer segmentation using data mining, Journal of Database Marketing and Customer Strategy Management, Vol. 19, No. 3, pp. 197-208, 2012.

### Implementation

Libraries

```{r message = FALSE, warning = FALSE}
library(stringr)
library(arules)
library(tidyverse)
library(cluster)
library(rpart)
library(rpart.plot)
library(recommenderlab)
```

Load and preprocess Online Retail Data

```{r}
load(url("http://vault.data-r.com/data/onlineretail.rda"))
dataset <- onlineretail

### Split Date and Time

Day <- as.Date(substr(dataset$InvoiceDate,1, 8), "%d/%m/%y")
Time <- substr(as.character(dataset$InvoiceDate), 10, 14)

dataset <- cbind(dataset, Day, Time)

### Keep only UK

pos_uk <- which(dataset$Country == "United Kingdom")
dataset <- dataset[pos_uk,]
```

Create RMF Dataset per CustomerID

```{r}
dataset <- as_tibble(dataset)
dataset$UnitPrice <- as.numeric(str_replace(dataset$UnitPrice, ",", "."))

RMFAnalysis <- summarise(group_by(dataset, CustomerID), 
                         Frequency = n(), 
                         Recency.date = max(Day),
                         Monetary.mean = mean(Quantity * UnitPrice, na.rm = TRUE), 
                         Monetary.sum = sum(Quantity * UnitPrice, na.rm = TRUE))

RMFAnalysis <- cbind(RMFAnalysis, max(RMFAnalysis$Recency.date) - RMFAnalysis$Recency.date)
names(RMFAnalysis)[ncol(RMFAnalysis)] <- "Recency"

### Remove some problematic Customers 

pos_problem <- c(which(RMFAnalysis$Monetary.mean <= 0), 
                 which(RMFAnalysis$Monetary.sum <= 0),
                 which(RMFAnalysis$Monetary.sum >= quantile(RMFAnalysis$Monetary.sum, 0.99)))

RMFAnalysis <- na.omit(RMFAnalysis[-pos_problem,])

RMFAnalysis.select <- RMFAnalysis[,c(2,5,6)]
RMFAnalysis.scale <- data.frame(as.numeric(scale(RMFAnalysis.select[,1])), 
                                as.numeric(scale(RMFAnalysis.select[,2])), 
                                as.numeric(scale(RMFAnalysis.select[,3])))  

# Correlations

cor(as.numeric(RMFAnalysis$Recency), RMFAnalysis$Frequency)
cor(as.numeric(RMFAnalysis$Frequency), RMFAnalysis$Monetary.mean)
```

Analysis for e.g. 4-6 clusters (customer groups)

```{r}
clusters <- 4

fit <- kmeans(RMFAnalysis.scale, clusters)
print(aggregate(RMFAnalysis.select, by=list(fit$cluster), FUN=mean))

RMFAnalysis <- data.frame(RMFAnalysis, fit$cluster)

for(cluster in 1:clusters) {
  cluster_data <- RMFAnalysis[which(RMFAnalysis$fit.cluster==cluster),]
  
  ### Classification / Decision Trees
  
  fit <- rpart(Monetary.sum ~ Frequency + Recency, data=cluster_data)
  rpart.plot(fit)

  ### Association Rules
  
  pos_transaction <- which(dataset$CustomerID %in%
                           RMFAnalysis$CustomerID[which(RMFAnalysis$fit.cluster==cluster)])
  write.csv2(dataset[pos_transaction,], file="transactions.txt", row.names = FALSE)
  trans <- read.transactions("transactions.txt", format="single", sep=";", cols=c(1,3), rm.duplicates=TRUE)
  
  support <- 0.03
  confidence <- 0.5
  
  rules <- apriori(trans, parameter = list(supp = support, conf = confidence, target = "rules"))
  if(length(rules) > 0) { inspect(head(rules, n = 30, by = "confidence")) }
}
```

User-based Collaborative Filtering

```{r}
onlineretail.clean <- na.omit(onlineretail)

products <- as(data.frame(onlineretail.clean[,c(7, 2)]), "binaryRatingMatrix")
image(products[1:20,1:20])

recommender <- Recommender(products, method="POPULAR")

# Create recommendations using user ids with ids 1..10 in the training data
pre <- predict(recommender, 1:10, data = products, n = 10)
as(pre, "list")
```
