---
title: Linear Regression - Used Car Prices
output: html_document
---

### About

This example is taken from the introductory lecture on statistics at the WU Vienna University of Economics and Business (PI Statistik). The main task is to build a (linear regression) model for used car prices of one specific model of a car given the prices of many used car retailers. Furthermore, the goal is to create a price prediction model based on a given set of attributes.

### Implementation

First we load all required libraries. We need the library `MASS` for the step-wise variable selection at the end of this script.

```{r message = FALSE, warning = FALSE}
library(MASS)
```

Furthermore, we load the sample data from our data repository online

```{r}
load(url("http://vault.data-r.com/data/usedcars.rda"))
```

and check descriptive statistics of the data

```{r}
dim(usedcars)
head(usedcars)
summary(usedcars)
```

So we see that we have 100 samples with five attributes. Clearly, `price` is the one we like to predict. Now we may proceed doing an univariate linear regression using the most likely attribute influencing the price, i.e. `miles`

```{r}
result <- lm(price ~ miles, data=usedcars)
summary(result)
```

and plot the results

```{r echo = FALSE}
plot(usedcars$miles, usedcars$price)
abline(coefficients(result))
```

Let's try another univariate regression using `service`

```{r}
result <- lm(price ~ service, data=usedcars)
summary(result)
```

and plot the results again

```{r echo = FALSE}
plot(usedcars$service, usedcars$price)
abline(coefficients(result))
```

Now, let's try a multi-variate regression using all attributes

```{r}
result <- lm(price ~ miles + service + age + color, data=usedcars)
summary(result)
```

And finally, apply a variable selection, i.e. step-wise selection (exact AIC)

```{r}
fit <- lm(price~miles+color+service+age,data=usedcars)
step <- stepAIC(fit, direction="both")
step$anova # display results 
```
