---
title: Wholesale in Portugal - Clustering
output: html_document
---

### About

This data set is taken from http://archive.ics.uci.edu/ml/datasets/Wholesale+customers and contains data of a wholesale distributor in Portugal, i.e. annual spending in monetary units on diverse product categories:

1) FRESH: annual spending on fresh products (Continuous)
2) MILK: annual spending on milk products (Continuous)
3) GROCERY: annual spending on grocery products (Continuous)
4) FROZEN: annual spending on frozen products (Continuous)
5) DETERGENTS_PAPER: annual spending on detergents and paper products (Continuous)
6) DELICATESSEN: annual spending on and delicatessen products (Continuous)
7) CHANNEL: Horeca (Hotel/Restaurant/Cafe) or Retail channel (Nominal)
8) REGION: Lisbon, Oporto or Other (Nominal)

### Implementation

First, load sample data from our data repository online

```{r}
load(url("http://vault.data-r.com/data/wholesale.rda"))
```

We only use the continuous spending variables

```{r}
data <- wholesale[,3:8]
```

#### Partitional (k-Means) Clustering

First, we plot a so-called scree plot to determine the amount of clusters

```{r}
max_clusters <- 15
wss <- (nrow(data)-1)*sum(apply(data,2,var))
for (i in 2:max_clusters) wss[i] <- sum(kmeans(data,centers=i)$withinss)
plot(1:max_clusters, wss, type="b", xlab="Number of Clusters", ylab="Within groups sum of squares") 
```

Then we apply a standard k-means clustering

```{r}
n_cluster <- 4
fit <- kmeans(data, n_cluster)

aggregate(data, by=list(fit$cluster), FUN=mean)
wholesale.cluster <- data.frame(wholesale, fit$cluster) 
names(wholesale.cluster)[ncol(wholesale.cluster)] <- "Cluster"
head(wholesale.cluster)
```

#### Hierachical Clustering

```{r}
d <- dist(data, method = "euclidean")
fit <- hclust(d, method="ward.D")
plot(fit)

n_cluster <- 4
groups <- cutree(fit, k=n_cluster)
rect.hclust(fit, k=n_cluster, border="red") 
```
